Sea: A lightweight data-placement library for Big Data scientific computing

### 2022-04-13 paper/sea-comp/paper.tex

\documentclass[10pt,journal,compsoc]{IEEEtran}
\usepackage[table]{xcolor}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{ulem}
\usepackage{graphicx}
\usepackage{rotating}
\usepackage{setspace}
\usepackage{longtable}
\usepackage[binary-units=true]{siunitx}
\usepackage{multirow}
\usepackage{subcaption}
%\usepackage{algorithm}
\usepackage[ruled,norelsize]{algorithm2e}
\usepackage{algpseudocode}
\usepackage{amsmath}
\usepackage{tabularx}

@@

\newcommand\VRule[1][\arrayrulewidth]{\vrule width #1}
% \algnewcommand\algorithmicforeach{\textbf{for each}}
% \algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
% \algblock{Input}{EndInput} \algnotext{EndInput} \algblock{Output}{EndOutput}
% \algnotext{EndOutput} \newcommand{\Desc}[2]{\State \makebox[2em][l]{#1}#2}
\newcommand{\todo}[1]{\marginpar{\parbox{18mm}{\flushleft\tiny\color{red}\textbf{TODO}:
#1}}}

@@

      Data locality ensures that compute tasks are scheduled nearest to where
      the data is located. Historically, storage and compute nodes were distinct
      layers, data transfer to the compute location. Rather than transferring
      large amounts of data over the network to compute tasks, which could incur
      significant overheads, Big Data frameworks ensure that data is stored
      directly on the compute nodes. When a compute task requires access to
      specific data, the scheduler sends the task to the nearest available node
      to the data, thereby minimizing any cost of network-related data
      transfers. This strategy is not only used by Big Data Frameworks such as
      Hadoop MapReduce, Apache Spark, and Dask~\cite{rocklin2015dask}, but also
      enabled by file systems such as HDFS and Alluxio.

@@

Glibc interception is achieved by writing wrappers to existing glibc functions.
Most importantly, every glibc function containing accepting a file path as input parameter needs to be
wrapped. The wrappers take any input filepath that is located within the
user-provided Sea mountpoint and convert it a filepath pointing to the best
available storage device.

@@

\subsubsection{Memory-management}
      \begin{table}
      \centering
      \begin{tabular}{ccc}
        \toprule
       Mode & .sea\_flushlist & .sea\_evictlist \\
       \midrule
       Copy & yes & no \\
       Remove & no & yes \\
       Move & yes & yes \\
       Keep & no & no \\

@@

      Experimental results can be prone to error and variability, particularly
      in instances where there a many factors at play, such as in distributed cluster
      environments. To improve confidence in our results, we developed a
      simplified performance model that could help describe the behaviours observed in our experimental
      results. Since different PFS may
      operate differently, our baseline model will be based on Lustre which is
      commonly used on HPC infrastructure.

@@

      As with many file systems, page cache plays an important role in the speed
      of application read and writes in Lustre. Since the effects of page cache
      may be non-negligible given amount of memory available and the data
      accessed during the execution of the application, it is important to
      include it in our model. The makespan of an application I/O to and from
      page cache, $M_{c}$, can be described as the following:
      Equation~\ref{eq:sea-comp:lustrenpc} where it is assumed that none of the
      data is written or read from page cache.

@@

      Here, $M_{n}$ represents the makespans of the different possible storage
      levels (e.g. tmpfs, NVMe, SSD, HDD, Lustre) and $M_{S}$ represent the
      makespan of an application reading and writing to Sea. For our model, we
      will assume 3 storage layers: 1) fast tmpfs, 2) intermediate local SSD
      storage, and 3) slow parallel file system layer. It is important to note that
      the model assumes that I/O never occurs in distinct storage layers in parallel.
      Although Sea tries to exhaust the layers, starting from the fastest, it is Nevertheless
      possible that Sea writes to two storage layers at the same time, for instance, when one becomes full
      and another parallel thread needs to perform a write.

@@

      \begin{equation}\label{eq:sea-comp:mst}
          M_{st} = \frac{D_{tr}}{cC_{r}} + \frac{D_{tw}}{cC_{w}}
      \end{equation}

@@

      To evaluate the Lustre and Sea models defined in
      Subsection~\ref{ss:sea-comp:model} and the real performance of both file
      systems with data intensive applications, we wrote a simple Python
      application based off of Algorithm~\ref{alg:sea-comp:incrementation}.

@@

      As can be seen in Figure~\ref{fig:sea-comp:benchmarks}, Sea using
      in-memory configuration speeds up execution the majority of the
      conditions. The largest speedups observed can be seen in
      Figure~\ref{fig:sea-comp:processes} at 32 processes, where the speedup is
      nearly 3$\times$. We believe that the speedup is primarily due to the
      contention on Lustre. While each local disk has an average of 5 processes
      trying to write concurrently to it at 32 processes, each Lustre has around
      3 concurrent processes while to it. Whereas one might assume this means
      Lustre would have less contention overall and should perform better, there
      are significant bandwidth differences to account for between the local
      disks and Lustre (see Table~\ref{table:sea-comp:fs}) that would result in
      improved local disk performance despite increased contention. Furthermore,
      Lustre has a centralized metadata server that is responsible for
      determining which OST is assigned to each block, guaranteeing a certain
      amount of load-balance on the storage disks. This is in contrast with the
      local disks, who do not rely on a metadata server and are selected by Sea
      via a random shuffling.

@@

    The flush-and-evict process may affect overall application parallelism. It
    would depend on how many Sea processes are started on the compute node. If
    only a single instance of Sea is called on a compute node, there will only
    be a single flush and evict process, and therefore, only a single flush and
    evict process will exist. However, if Sea is launched many times on a given
    node, there will be many flush and evict processes which may interfere with
    the application compute. However, our results do not indicate that
    performance is significantly impacted by the presence of a single flush and
    evict process.

@@

    On shared cluster environments, such as though found in high-performance
    computing clusters, it is less likely that the PFS would be so underused
    that it could achieve better performance than leveraging local storage, even
    if local storage was limited. We therefore believe that it is very unlikely
    obtain no speedups from using Sea on a production HPC cluster as long as the
    application is data intensive.

@@

\appendices
% you can choose not to have a title for an appendix if you want by leaving the
% argument blank
\section{}
Appendix two text goes here.

% use section* for acknowledgment
\ifCLASSOPTIONcompsoc
  % The Computer Society usually uses the plural form
  \section*{Acknowledgments}
\else
  % regular IEEE prefers the singular form
  \section*{Acknowledgment}
\fi

>>>>

\documentclass[10pt,journal,compsoc]{IEEEtran}
\usepackage[table]{xcolor}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{ulem}
\usepackage{graphicx}
\usepackage{rotating}
\usepackage{setspace}
\usepackage{longtable}
\usepackage[binary-units=true]{siunitx}
\usepackage{multirow}
\usepackage{subcaption}
%\usepackage{algorithm}
\usepackage[ruled,norelsize]{algorithm2e}
\usepackage{algpseudocode}
\usepackage{amsmath}
\usepackage{tabularx}
\usepackage{xspace}

@@

\newcommand\VRule[1][\arrayrulewidth]{\vrule width #1}
% \algnewcommand\algorithmicforeach{\textbf{for each}}
% \algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
% \algblock{Input}{EndInput} \algnotext{EndInput} \algblock{Output}{EndOutput}
% \algnotext{EndOutput} \newcommand{\Desc}[2]{\State \makebox[2em][l]{#1}#2}
\newcommand{\todo}[1]{\marginpar{\parbox{18mm}{\flushleft\tiny\color{red}\textbf{TODO}:
#1}}}
\newcommand{\TG}[1]{\color{red}\textsc{From Tristan}: #1\xspace\color{black}}

@@

      Data locality ensures that compute tasks are scheduled nearest to where
      the data is located. Historically, storage and compute nodes were distinct
      layers, \TG{missing word?} data transfer to the compute location. Rather than transferring
      large amounts of data over the network to compute tasks, which could incur
      significant overheads, Big Data frameworks ensure that data is stored
      directly on the compute nodes. When a compute task requires access to
      specific data, the scheduler sends the task to the nearest available node
      to the data, thereby minimizing any cost of network-related data
      transfers. This strategy is not only used by Big Data Frameworks such as
      Hadoop MapReduce, Apache Spark, and Dask~\cite{rocklin2015dask}, but also
      enabled by file systems such as HDFS and Alluxio.

@@

Glibc interception is achieved by writing wrappers to existing glibc functions.
Most importantly, every glibc function accepting a file path as input parameter needs to be
wrapped. The wrappers take any input filepath that is located within the
user-provided Sea mountpoint and convert it to a filepath pointing to the best
available storage device.

@@

\subsection{Memory-management}
      \begin{table}
      \centering
      \begin{tabular}{ccc}
        \toprule
       Mode & .sea\_flushlist & .sea\_evictlist \\
       \midrule
       Copy & yes & no \\
       Remove & no & yes \\
       Move & yes & yes \\
       Keep & no & no \\

@@

      Experimental results can be prone to error and variability, particularly
      in instances where there a many factors at play, such as in distributed cluster
      environments. To improve confidence in our experimental results, we developed a
      simplified performance model that could help describe the behaviours observed in our experimental
      results. Since different PFS may
      operate differently, our baseline model will be based on Lustre which is
      commonly used on HPC infrastructure.

@@

      As with many file systems, page cache plays an important role in the speed
      of application read and writes in Lustre. Since the effects of page cache
      may be non-negligible given amount of memory available and the data
      accessed during the execution of the application, it is important to
      include it in our model. The makespan of an application I/O to and from
      page cache, $M_{c}$, can be described as in
      Equation~\ref{eq:sea-comp:lustrenpc}, where it is assumed that none of the
      data is written or read from page cache.

@@

      Here, $M_{n}$ represents the makespans of the different possible storage
      levels (e.g. tmpfs, NVMe, SSD, HDD, Lustre) and $M_{S}$ represent the
      makespan of an application reading and writing to Sea. For our model, we
      will assume 3 storage layers: 1) fast tmpfs, 2) intermediate local SSD
      storage, and 3) slow parallel file system layer. It is important to note that
      the model assumes that I/O never occurs in distinct storage layers in parallel.
      Although Sea tries to exhaust the layers, starting from the fastest, it is nevertheless
      possible that Sea writes to two storage layers at the same time, for instance, when one becomes full
      and another parallel thread needs to perform a write.

@@

      \begin{equation}\label{eq:sea-comp:mst}
          M_{St} = \frac{D_{tr}}{cC_{r}} + \frac{D_{tw}}{cC_{w}}
      \end{equation}

@@

      To evaluate the performance of Sea with data intensive applications, we wrote a simple Python
      application based off of Algorithm~\ref{alg:sea-comp:incrementation}.

@@

      As can be seen in Figure~\ref{fig:sea-comp:benchmarks}, Sea using
      in-memory configuration speeds up execution the majority of the
      conditions. The largest speedups observed can be seen in
      Figure~\ref{fig:sea-comp:processes} at 32 processes, where the speedup is
      nearly 3$\times$. We believe that the speedup is primarily due to the
      contention on Lustre. While each local disk has an average of 5 processes
      trying to write concurrently to it at 32 processes, each Lustre has around
      3 concurrent processes while to it \TG{this sentence doesn't make sense, revise}. Whereas one might assume this means
      Lustre would have less contention overall and should perform better, there
      are significant bandwidth differences to account for between the local
      disks and Lustre (see Table~\ref{table:sea-comp:fs}) that would result in
      improved local disk performance despite increased contention. Furthermore,
      Lustre has a centralized metadata server that is responsible for
      determining which OST is assigned to each block, guaranteeing a certain
      amount of load-balance on the storage disks. This is in contrast with the
      local disks, who do not rely on a metadata server and are selected by Sea
      via a random shuffling.

@@

    The flush-and-evict process may affect overall application parallelism. It
    would depend on how many Sea processes are started on the compute node. If
    only a single instance of Sea is called on a compute node, there will only
    be a single flush and evict process, and therefore, only a single flush and
    evict process will exist \TG{this sentence is circular, revise}. However, if Sea is launched many times on a given
    node, there will be many flush and evict processes which may interfere with
    the application compute. However, our results do not indicate that
    performance is significantly impacted by the presence of a single flush and
    evict process.

@@

    On shared cluster environments, such as those found in high-performance
    computing clusters, it is less likely that the PFS would be so underused
    that it could achieve better performance than leveraging local storage, even
    if local storage was limited. We therefore believe that it is very unlikely
    obtain no speedups from using Sea on a production HPC cluster as long as the
    application is data intensive.

@@

% \appendices
% % you can choose not to have a title for an appendix if you want by leaving the
% % argument blank
% \section{}
% Appendix two text goes here.

% % use section* for acknowledgment
% \ifCLASSOPTIONcompsoc
%   % The Computer Society usually uses the plural form
%   \section*{Acknowledgments}
% \else
%   % regular IEEE prefers the singular form
%   \section*{Acknowledgment}
% \fi
